---
title: "Night vs Day Lapse Rate"
author: "Danielle Reimanis"
date: "12/5/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyverse)
library(tidyr)
library(lubridate)
library(ggthemes)
library(ggplot2)
library(darksky)
library(purrr)

```

```{r, include=FALSE}
#load('./all_sites.Rdata')
```

## Pulling Sunrise and Sunset Data

```{r}
latitude <- 40.556899
longitude <- -105.611251

starting_daily <- get_forecast_for(latitude, longitude, "2019-07-04T01:00:00",
                                   units = "us", language = "en",
                                  exclude = 'currently, hourly, minutely',
                                  add_json = FALSE, add_headers =FALSE)
combined_daily <- starting_daily$daily%>%
  select(1,4,5)

date_vector <- seq(from = as.Date("2019-07-05"), to = as.Date("2019-10-13"), by = 1)

for(i in as.list(date_vector)){
    print(i)
  a_timestamp <- paste(i, "T01:00:00", sep = "")
  b_forecast <- get_forecast_for(latitude, longitude, a_timestamp, 
                                 units = "us", language = "en",
                                 exclude = 'currently, hourly, minutely',
                                 add_json = FALSE, add_headers =FALSE)
  c_forecast <- b_forecast$daily%>%
    select(1,4,5)

  combined_daily <- rbind(combined_daily, c_forecast)
}

```

## Merging Sunrise/Sunset data with temperature data

```{r}
temp_data <- merge_temp_location %>%
  rename(datetime = date)%>%
  mutate(date = date(datetime))

combined_daily_for_merge <- combined_daily%>%
  mutate(date = seq(from = as.Date("2019-07-04"), 
                    to = as.Date("2019-10-13"), by = 1))

temp_sun_data <- merge(temp_data, combined_daily_for_merge, by = 'date')

```

## Data organized by day and night & averaged

```{r}

day_and_night <- temp_sun_data %>%
  select(-time)%>%
  mutate(day_night = ifelse(datetime > sunriseTime & datetime < sunsetTime,
                            'day', 'night'))%>%
  arrange(datetime)

day_night_average <- day_and_night %>%
  group_by(site, date, day_night)%>%
  summarise(avg_temp = mean(temp),
            elev = mean(elev),
            utmE = mean(utmE),
            utmN = mean(utmN))%>%
  arrange(date)

```

## Plotting Day vs Night Averages

```{r}

ggplot(day_night_average, aes(x = date, y= avg_temp, 
                              group = site, color = day_night))+
  geom_point()+
  theme_linedraw()+
  labs(x='Date', y= 'Average Temp (°C)')

ggplot(day_night_average, aes(x=date, y=avg_temp,
                                 color = day_night))+
  geom_point()+
  facet_wrap(~site)+
  theme_linedraw()+
  labs(x='Date', y= 'Average Temp (°C)')

```

## Regression

```{r}

day_night_wide <- day_night_average %>%
  spread(., key = 'day_night',
         value = 'avg_temp')%>%
  mutate(doy = yday(date))%>%
  arrange(doy)%>%
  mutate(elev_km = elev/1000)

ydays <- day_night_wide %>%
  mutate(doy = yday(date))%>%
  pull(doy)%>%
  unique()

day_lm_fun <- function(x){
 filter_temp <- day_night_wide%>%
    filter(doy == x)
 
  daily_lapse_lm <- lm(day~elev_km, data = filter_temp)
}

day_lm <- ydays%>%
  map(~day_lm_fun(.x))%>%
  set_names(ydays)

day_lm_slopes <- day_lm%>%
  map(~coef(.x)[2])%>%
  simplify()%>%
  tibble()%>%
  rename(Day = 1)%>%
  mutate(date = seq(from = as.Date("2019-07-04"), 
                    to = as.Date("2019-10-13"), by = 1))


night_lm_fun <- function(x){
 filter_temp <- day_night_wide%>%
    filter(doy == x)
 
  daily_lapse_lm <- lm(night ~ elev_km, data = filter_temp)
}

night_lm <- ydays%>%
  map(~night_lm_fun(.x))%>%
  set_names(ydays)

night_lm_slopes <- night_lm%>%
  map(~coef(.x)[2])%>%
  simplify()%>%
  tibble()%>%
  rename(Night = 1)%>%
  mutate(date = seq(from = as.Date("2019-07-04"), 
                    to = as.Date("2019-10-13"), by = 1))

```

## Joining Night and Day Linear Models and Plotting

```{r}

night_day_lm <- merge(day_lm_slopes, night_lm_slopes, by = 'date')

night_day_slopes_long <- night_day_lm %>%
  gather(., key = 'Time',
         value = 'lapse_rate',
         -date)

ggplot(night_day_slopes_long, aes(x=date, y=lapse_rate, color = Time))+
  geom_point()+
  theme_linedraw()+
  theme(legend.position = c(0.75, 0.75))+
  labs(x = 'Date', y = 'Lapse Rate (°C/km)')

ggplot(night_day_slopes_long, aes(x=date, y=lapse_rate, color = Time))+
  geom_line()+
  theme_linedraw()+
  theme(legend.position = c(0.75, 0.75))+
  labs(x = 'Date', y = 'Lapse Rate (°C/km)')

```

## Zoomed in, while all iButtons were 'online'

```{r}

online_night_day_slope <- night_day_slopes_long%>%
  filter(date > "2019-07-13")

ggplot(online_night_day_slope, aes(x=date, y=lapse_rate, color = Time))+
  geom_line()+
  theme_linedraw()+
  theme(legend.position = c(0.6, 0.85))+
  labs(x = 'Date', y = 'Lapse Rate (°C/km)')

```

## Pulling R^2 Values

```{r}
day_lm_sum_r <- day_lm %>%
  map(~summary(.x)$r.squared)%>%
  tibble()%>%
  rename(Day = 1)%>%
  mutate(date = seq(from = as.Date("2019-07-04"), 
                    to = as.Date("2019-10-13"), by = 1))


night_lm_sum_r <- night_lm %>%
  map(~summary(.x)$r.squared)%>%
  tibble()%>%
  rename(Night = 1)%>%
  mutate(date = seq(from = as.Date("2019-07-04"), 
                    to = as.Date("2019-10-13"), by = 1))

day_night_r <- merge(day_lm_sum_r, night_lm_sum_r, by = 'date')

day_night_r_long <- day_night_r %>%
  gather(., key = 'Time',
         value = 'R_squared',
         -date)

r_plotting_ready <- day_night_r_long %>%
  mutate(R_squared = as.numeric(as.list(R_squared)))%>%
  rename(dateR = 1, TimeR = 2)

```

## Plotting R^2 Values and creating data frame with Slope and R Squared

```{r}

ggplot(r_plotting_ready, aes(x=dateR, y= R_squared, color = TimeR))+
  geom_point()+
  theme_linedraw()+
  theme(legend.position = c(0.075, 0.11), legend.title=element_blank())+
  labs(x = 'Date', y = 'R Squared')

slope_R_all <- cbind(r_plotting_ready, night_day_slopes_long)%>%
  select(3,4:6)

```

## Plotting R squared values and slopes

```{r}

ggplot(slope_R_all, aes(x=date, y=lapse_rate,
                        color=Time, size=R_squared))+
  geom_point(alpha=0.67)+
  theme_linedraw()+
  labs(x = 'Date', y = 'Lapse Rate (°C/km)')

```


